NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids

### 2020-01-03 Manuscript/manuscriptPGPE.tex
Add figure from issue https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/9 into the manuscript

Without further discussion about poorly performing force fields, we focus on more detailed analysis
of CHARMM36 simulations, which give the best overall agreement with experiments for
the headgroup and glycerol backbone order parameters of all PC, PS, PG and PE lipids in this and our previous
studies \cite{botan15,antila19}. Even thought many order parameters in CHARMM36 simulations are
not within the experimental error bars, it seems to capture the essential differences between different headroups.
%(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}).
In previous study, CHARMM36 predicted the more negative $\beta$-carbon order parameter and larger forking of
the $\alpha$-carbon in PS headgroup than in PC \cite{antila19}.
In this work,
%the CHARMM36 simulations reproduce also the other essential differences in experimental
%headgroup order parameters between different headgroups (Fig. \ref{HGorderParameters}) despite the inaccuracies
%in individual segments:
the PE headgroup order parameters are similar to the PC \cite{botan15}
and $\beta$-carbon order parameters is positive in PG headgroup, while negative values observed in
other lipids. Therefore, we use CHARMM36 simulations to analyze the structural differences between
headgroups (Fig.~\ref{structures}).
\todo{The structures are already in Fig. \ref{structures}, but we need also the dihedral distributions to finish the discussion.
Maybe we should analyze also some other structural features, at least P-N angles. More discussion in issue
https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/9}.
%While rotation around N-C$_\alpha$-C$_\beta$-O$_\alpha$ dihedral was
%significantly different in PS and PC headgroups in previous work \cite{antila19}, essential differences
%between PC, PE and PG are not observed here (Fig.~\ref{structures}) \todo{We need also the dihedral distributions to finish this discussion.}. \\
%\todo{Why is the $\beta$-carbon order parameter of PG different to PC and PE then?}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPG.pdf}
  \caption{\label{HGorderParametersPOPG}
    The headgroup and glycerol backbone order parameters of PG lipids
    from experiments (POPG and signs from this work and from Ref.~\citenum{borle85}, %contains 10mM of PIPES,
    DPPG with 100mM NaCl from Ref.~\citenum{wohlgemuth80},% contains 10mM PIPES and,
    and E.Coli PG results from Ref.~\citenum{gally81}).
    and simulations with different force fields.
  }
\end{figure}
\begin{figure*}[!h]
  \centering
  \includegraphics[width=18.0cm]{../Figs/struct.png}
  \caption{\label{structures}
    Overlayed snapshots from CHARMM36 simulations of different lipids which give the best agreement with experiments.
  }
\end{figure*}

>>>>

Without further discussion about poorly performing force fields, we focus on more detailed analysis
of CHARMM36 simulations, which give the best overall agreement with experiments for
the headgroup and glycerol backbone order parameters of all PC, PS, PG and PE lipids in this and our previous
studies \cite{botan15,antila19}. Even thought many order parameters in CHARMM36 simulations are
not within the experimental error bars, it seems to capture the essential differences between different headroups.
%(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}).
In previous study, CHARMM36 predicted the more negative $\beta$-carbon order parameter and larger forking of
the $\alpha$-carbon in PS headgroup than in PC \cite{antila19}.
In this work,
%the CHARMM36 simulations reproduce also the other essential differences in experimental
%headgroup order parameters between different headgroups (Fig. \ref{HGorderParameters}) despite the inaccuracies
%in individual segments:
the PE headgroup order parameters are similar to the PC \cite{botan15}
and $\beta$-carbon order parameters is positive in PG headgroup, while negative values observed in
other lipids. Therefore, we use CHARMM36 simulations to analyze the structural differences between
headgroups (Fig.~\ref{structures}).
\todo{The structures and dihedrals are in Fig. \ref{structures}, but we need to finish the discussion.
Maybe we should analyze also some other structural features, at least P-N angles. More discussion in issue
https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/9}.
%While rotation around N-C$_\alpha$-C$_\beta$-O$_\alpha$ dihedral was
%significantly different in PS and PC headgroups in previous work \cite{antila19}, essential differences
%between PC, PE and PG are not observed here (Fig.~\ref{structures}) \todo{We need also the dihedral distributions to finish this discussion.}. \\
%\todo{Why is the $\beta$-carbon order parameter of PG different to PC and PE then?}

@@

\begin{figure}[!h]
  \centering
  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPG.pdf}
  \caption{\label{HGorderParametersPOPG}
    The headgroup and glycerol backbone order parameters of PG lipids
    from experiments (POPG and signs from this work and from Ref.~\citenum{borle85}, %contains 10mM of PIPES,
    DPPG with 100mM NaCl from Ref.~\citenum{wohlgemuth80},% contains 10mM PIPES and,
    and E.Coli PG results from Ref.~\citenum{gally81}).
    and simulations with different force fields.
  }
\end{figure}
\begin{figure*}[!h]
  \centering
  \includegraphics[width=18.0cm]{../Figs/PCPEPGPSstructures.png}
  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals1.png}
  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals2.png}
  \caption{\label{structures}
    Overlayed snapshots and dihedral angle distributions from CHARMM36 simulations of different lipids which give the best agreement with experiments.
  }
\end{figure*}
